Unrecognized States: Theory and Cases

### 2017-07-04 theory.tex
tighten up section 1

Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders. In practice we often observe groups of states like the OECD or the UN acting in this capacity. We also assume player $c$ prefers peace to war; this implies that player $c$ will not fund a military buildup that it expects will induce war.

In contrast, Player $p$ most prefers recognized independence and opposed independence, aligning its interests with the secessionists. We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium. Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan et al., 2012), e.g. as Russia does to Georgia via South Ossetia and Abkhazia; 2) ethnic solidarity with the secessionists (e.g. Turkey's support of the Turkish Republic of Northern Cyprus); 3) hope of eventual annexation of the disputed territory (e.g. Armenia's support of Nagorno-Karabakh). Although there may exist patrons whose most-preferred outcome is the status quo, we examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. We will show that even in this circumstance the status quo remains an equilibrium outcome.

@@

All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment. The status quo payoffs for the secessionists are automatically reduced by $\mu$ each period. The steady reduction by $\mu$ represents the costs of non-recognition.

Only two outcomes of the stage game do not lead to absorbing states. If both $s$ and $g$ play Status Quo, then the status quo persists. Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other.

If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game ends with payoffs in every subsequent period given by the corresponding payoffs in the stage game. Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.

There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously. We use a lottery to determine whether the secessionists or government wins the war. Victory would, among other things, allow an unrecognized state to force recognition by the home state government.

Future payoffs are discounted with a common parameter $\de \in [0,1]$. Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.

The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players.

>>>>

Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders.\footnote{In practice we often observe groups of states like the OECD or the UN acting in this capacity.} We also assume player $c$ prefers peace to war; this implies that player $c$ will not fund a military buildup that it expects will induce war.

In contrast, Player $p$ most prefers recognized independence and opposes independence, aligning its interests with the secessionists. We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium. Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan et al., 2012), e.g. as Russia does to Georgia via South Ossetia and Abkhazia; 2) ethnic solidarity with the secessionists (e.g. Turkey's support of the Turkish Republic of Northern Cyprus); 3) hope of eventual annexation of the disputed territory (e.g. Armenia's support of Nagorno-Karabakh). Although there may exist patrons whose most-preferred outcome is the status quo, we examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. We will show that even in this circumstance the status quo remains an equilibrium outcome.

@@

All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment. The status quo payoffs for the secessionists are automatically reduced by $\mu$ each period, reflecting the costs of non-recognition.

If both $s$ and $g$ play Status Quo, then the status quo persists. Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other. These are the only outcomes of the stage game that do not lead to absorbing states.

If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game ends with payoffs in every subsequent period given by the corresponding payoffs in the stage game---i.e., the result is a negotiated settlement benefiting the player who did not cede.

If either of the parties attacks first or both attack simultaneously, the result is war. We use a lottery to determine whether the secessionists or government wins the war. The victor is able to force recognition/reunification.

Future payoffs are discounted with a common parameter $\de \in [0,1]$. The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players.
